---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mirageV2

<!-- badges: start -->
[![R-CMD-check](https://github.com/anson-li8/mirageV2/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/anson-li8/mirageV2/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/anson-li8/mirageV2/graph/badge.svg)](https://app.codecov.io/gh/anson-li8/mirageV2)
[![test-coverage](https://github.com/anson-li8/mirageV2/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/anson-li8/mirageV2/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

`mirageV2` is a refactor and performance improvement of the
[`mirage`](https://github.com/xinhe-lab/mirage) package for rare variant
(RV) association testing using a Bayesian mixture model. It models variants
in a gene as a mixture of risk and non-risk variants, with prior
probabilities informed by functional annotations such as conservation
scores and protein impact predictions.

## Key improvements over `mirage`

| Area | `mirage` | `mirageV2` |
|------|----------|------------|
| **Numerical stability** | Probability-space EM with `ifelse(bb==Inf, 3e300, bb)` overflow patches | Log-space EM with `log_sum_exp` throughout |
| **BF computation** | R `integrate()` per variant in a loop | Vectorized C++ composite Simpson's rule via Rcpp |
| **EM engine** | Duplicated `vapply`/nested loops in `mirage()` and `mirage_vs()` | Single unified engine, vectorized with `rowsum`/`table` |
| **Input validation** | Implicit column guessing | Strict S3 validation with `cli` error messages |
| **Output** | Loose unnamed lists | S3 classes (`mirage_result`, `mirage_vs_result`) with `print`, `summary`, `tidy` methods |
| **Testing** | None | 56 `testthat` tests covering edge cases |
| **Performance** | ~160s on `mirage_toy` (282 genes) | **~10s** (15x speedup) |

## Installation

```{r, eval = FALSE}
# Install devtools if needed
install.packages("devtools")

# Install mirageV2
devtools::install_github("anson-li8/mirageV2")
```

## Quick start

```{r example, eval = FALSE}
library(mirageV2)

# Load bundled toy data
toy <- read.csv(system.file("extdata", "mirage_toy.csv", package = "mirageV2"),
                stringsAsFactors = FALSE)

# Gene-level analysis
res <- mirage(toy, n1 = 4315, n2 = 4315)
print(res)
summary(res)

# Tidy output (broom-compatible)
tidy(res)

# Variant-set analysis
vs_toy <- read.csv(system.file("extdata", "mirage_vs_toy.csv", package = "mirageV2"),
                   stringsAsFactors = FALSE)
res_vs <- mirage_vs(vs_toy, n1 = 4315, n2 = 4315)
print(res_vs)
```

## Input format

### Gene-level (`mirage`)

A data frame with 5 columns (or 4 without `ID`):

| Column | Description |
|--------|-------------|
| `ID` | Variant identifier |
| `Gene` | Gene name |
| `No.case` | Allele count in cases |
| `No.contr` | Allele count in controls |
| `category` | Variant annotation group (or `group.index` for backward compatibility) |

### Variant-set level (`mirage_vs`)

Same as above without the `Gene` column.

## Method

`mirageV2` implements the same Bayesian mixture model as the original
`mirage`. For a gene $i$ with variants $j = 1, \ldots, m_i$:

- Each variant is modeled as risk ($Z_{ij} = 1$) or non-risk ($Z_{ij} = 0$)
  with $P(Z_{ij} = 1 \mid U_i = 1) = \eta_k$, where $k$ is the variant's
  annotation category.
- Each gene is a risk gene ($U_i = 1$) with probability $\delta$.
- Variant-level Bayes factors are computed by integrating over a
  Gamma-distributed relative risk parameter.
- Parameters $(\delta, \eta_1, \ldots, \eta_K)$ are estimated via EM in
  log-space.
- Gene-level posterior probabilities and likelihood ratio test p-values are
  reported.

See the [original `mirage`
documentation](https://xinhe-lab.github.io/mirage/) for full method
details.

## Session information

```{r}
sessionInfo()
```
