---
title: "Performance Benchmark: mirageV2 vs mirage"
author: "Anson Li"
output: rmarkdown::html_vignette
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

`mirageV2` is a performance-oriented rewrite of the original
[`mirage`](https://github.com/xinhe-lab/mirage) package. This vignette
benchmarks both packages on the same dataset and compares results for
numerical equivalence.

### Architecture differences

| Component | `mirage` | `mirageV2` |
|-----------|----------|------------|
| Bayes factor integration | R `integrate()` per variant in a loop | C++ composite Simpson's (vectorized) |
| EM algorithm | R `vapply`/nested loops, probability space | C++ EM, log-space with `log_sum_exp` |
| Numerical overflow handling | `ifelse(bb==Inf, 3e300, bb)` patch | Log-space arithmetic (no overflow possible) |

## Setup

```{r load}
library(mirageV2)

toy <- read.csv(
  system.file("extdata", "mirage_toy.csv", package = "mirageV2"),
  stringsAsFactors = FALSE
)
cat("Dataset:", nrow(toy), "variants across",
    length(unique(toy$Gene)), "genes\n")
```

Check if the original `mirage` package is available:

```{r check-original}
has_original <- requireNamespace("mirage", quietly = TRUE)
cat("Original mirage installed:", has_original, "\n")
```

## Runtime comparison

```{r benchmark, eval=has_original}
# Time mirageV2
t_v2 <- system.time(
  res_v2 <- mirage(toy, n1 = 4315, n2 = 4315, verbose = FALSE)
)

# Time original mirage
t_v1 <- system.time(
  res_v1 <- mirage::mirage(toy, n1 = 4315, n2 = 4315, verbose = FALSE)
)

cat("\n--- Timing Results ---\n")
cat(sprintf("mirage  (original): %.2fs\n", t_v1["elapsed"]))
cat(sprintf("mirageV2:           %.2fs\n", t_v2["elapsed"]))
cat(sprintf("Speedup:            %.1fx\n", t_v1["elapsed"] / t_v2["elapsed"]))
```

```{r benchmark-v2-only, eval=!has_original}
# Original mirage not installed; benchmark mirageV2 alone
t_v2 <- system.time(
  res_v2 <- mirage(toy, n1 = 4315, n2 = 4315, verbose = FALSE)
)
cat(sprintf("mirageV2: %.2fs\n", t_v2["elapsed"]))
cat("Install the original package for a direct comparison:\n")
cat("  devtools::install_github('xinhe-lab/mirage')\n")
```

## Result equivalence

```{r equivalence, eval=has_original}
# Compare posterior probabilities
v2_pp <- res_v2$BF.PP.gene$post.prob
v1_pp <- res_v1$BF.PP.gene[, 3]  # post.prob column

cat("Max absolute difference in posterior probabilities:",
    max(abs(v2_pp - v1_pp)), "\n")

# Compare delta estimates
cat(sprintf("mirage  delta: %.6f\n", res_v1$delta.est$delta.est))
cat(sprintf("mirageV2 delta: %.6f\n", res_v2$delta.est$delta.est))

# Compare eta estimates
cat("\nEta estimates (mirage vs mirageV2):\n")
print(data.frame(
  category = res_v2$eta.est |> rownames(),
  mirage = res_v1$eta.est$eta.est,
  mirageV2 = res_v2$eta.est$eta.est
))
```

## Session info

```{r}
sessionInfo()
```
