---
title: "Getting Started with mirageV2"
author: "Anson Li"
output: rmarkdown::html_vignette
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Installation

```{r, eval=FALSE}
devtools::install_github("anson-li8/mirageV2")
```

## Input format

`mirageV2` requires a data frame with variant-level summary statistics.

### Gene-level (`mirage`)

| Column | Description |
|--------|-------------|
| `ID` | Variant identifier |
| `Gene` | Gene name |
| `No.case` | Allele count in cases |
| `No.contr` | Allele count in controls |
| `category` | Variant annotation group (or `group.index`) |

### Variant-set level (`mirage_vs`)

Same as above without the `Gene` column.

## Gene-level analysis

```{r gene-example}
library(mirageV2)
library(generics)

toy <- read.csv(
  system.file("extdata", "mirage_toy.csv", package = "mirageV2"),
  stringsAsFactors = FALSE
)
head(toy)
```

Run the analysis:

```{r gene-run}
res <- mirage(toy, n1 = 4315, n2 = 4315)
```

The result is a `mirage_result` object:

```{r gene-print}
print(res)
```

Key outputs:

- `res$delta.est`: Estimated proportion of risk genes with p-value.
- `res$eta.est`: Estimated proportion of risk variants per category with p-values.
- `res$BF.PP.gene`: Bayes factors and posterior probabilities per gene.
- `res$BF.all`: Variant-level Bayes factors grouped by gene.

```{r gene-summary}
summary(res)
```

Tidy output:

```{r gene-tidy}
head(tidy(res))
```

## Variant-set analysis

```{r vs-example}
vs_toy <- read.csv(
  system.file("extdata", "mirage_vs_toy.csv", package = "mirageV2"),
  stringsAsFactors = FALSE
)
head(vs_toy)
```

```{r vs-run}
res_vs <- mirage_vs(vs_toy, n1 = 4315, n2 = 4315)
print(res_vs)
```

## Fixed eta

When external annotation data provides reliable risk proportions, fix `eta`
instead of estimating it:

```{r fixed-eta}
res_fixed <- mirage(toy, n1 = 4315, n2 = 4315,
                    estimate.eta = FALSE,
                    fixed_eta = c(0.60, 0.31, 0.16, 0.40, 0.20))
head(res_fixed$BF.PP.gene)
```

## Key parameters

| Parameter | Default | Description |
|-----------|---------|-------------|
| `gamma` | 3 | Gamma prior shape for effect size |
| `sigma` | 2 | Gamma prior scale for effect size |
| `eta.init` | 0.1 | Initial risk variant proportion |
| `delta.init` | 0.1 | Initial risk gene proportion |
| `max.iter` | 10000 | Maximum EM iterations |
| `tol` | 1e-5 | Convergence tolerance |

## Session info

```{r}
sessionInfo()
```
